# Results

```{r}
#| include: false

library(magrittr)
load(here:here("figures", "strata_table.Rda"))

```

The final data set used for this analysis consisted of 11,340
observations. All observations contained a TSH and Free T4 result and
less than three missing results from all other analytes selected for the
study. The dataset was then randomly split into a training set
containing 9071 observations and a testing set containing 2269
observations. The data was split using stratification of the Free T4
laboratory diagnostic value. @tbl-strata shows the split percentages.

```{r}
#| label: tbl-strata
#| tbl-cap: Data Stratification 
#| echo: false

strata_table %>% knitr::kable()

```

First, the report shows the ability of classification algorithms to
predict whether Free T4 will be diagnostic, with the prediction quality
measured by Area Under Curve (AUC) and accuracy. Data regarding the
importance association between each predictor analyte and the Free T4
Diagnostic value is then presented. Finally, data is presented with the
extent to which FT4 can be predicted by examining the correlation
statistics denoting the relationship between measured and predicted Free
T4 values.

## Predictability of Free T4 Classifications

In clinical decision-making, a key consideration in interpreting
numerical laboratory results is often just whether the results fall
within the normal reference range [@luo2016]. In the case of Free T4
reflex testing, the results will either fall within the normal range
indicating the Free T4 is not diagnostic of Hyper or Hypo Throydism, or
they will fall outside those ranges indicating they are diagnostic. The
final model achieved an accuracy of 0.796 and an AUC of 0.918.
@fig-roc_curve provides ROC curves for each of the four outcome classes.

![ROC curves for each of the four outcome
classes](figures/roc_curve_class){#fig-roc_curve}

@fig-conf-matrix-class shows the confusion matrix of the final testing
data. Of the 2269 total results, 1805 were predicted correctly, leaving
464 incorrectly predicted results. Of the incorrectly predicted results,
72 results predicted a diagnostic Free T4 when the correct result was
non-diagnostic. 392 of the incorrectly predicted results were predicted
as non-diagnostic when the correct result was diagnostic.

![Final Model Confusion
Matrix](figures/conf_matrix_class){#fig-conf-matrix-class}

## Contributions of Individual Analytes

Understanding how an ML model makes predictions helps build trust in the
model and is the fundamental idea of the emerging field of interpretable
machine learning (IML) [@greenwell2020]. @fig-vip-class shows the
importance of features in the final model. Importance can be defined as
the extent to which a feature has a \"meaningful\" impact on the
predicted outcome [@laan2006]. As expected, TSH is the leading variable
in importance rankings, leading all other variables by over 2000's
points. The following three variables are all parts of a Complete Blood
Count (CBC), followed by the patient's glucose value.

![Variable Importance Plot](figures/vip_class){#fig-vip-class}

## Predictability of Free T4 Results (Regression) 

Today, it has become widely accepted that a more sound approach to
assessing model performance is to assess the predictive accuracy via
loss functions. Loss functions are metrics that compare the predicted
values to the actual value (the output of a loss function is often
referred to as the error or pseudo residual) [@boehmke2020]. The loss
function used to evaluate the final model was selected as the Root Mean
Square Error, and the final testing data achieved an RMSE of 0.334.
@fig-reg-pred shows the plotted results. The predicted results were also
used to add the diagnostic classification of Free T4. These results
achieved an accuracy of 0.790, and thus very similar to the
classification model.

![Regression Predictions Plot](figures/reggression_pred){#fig-reg-pred}
